cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Updated cone data generation section

This function generates `n` points in a `p`-dimensional space, where the last dimension (`X_p`) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.

Points along the height axis are drawn from an exponential distribution to increase density near the tip:

$$
X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
$$

The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter:

$$
r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
$$

@@

The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$. This classical non-orientable surface loops back on itself with a half-twist.

@@

To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension. 
The first three dimensions form the classic $3\text{-}D$ Swiss roll shape. The $X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)$. For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

>>>>

This function generates $n$ points in a $p\text{-}D$, where the last dimension ($X_p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip. Points along the height axis are drawn from an exponential distribution to increase density near the tip. The $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$. The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter $r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1$.

@@

The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$ structure. This classical non-orientable surface loops back on itself with a half-twist.

@@

To further generalize the Swiss roll structure and introduce realistic noise, we define a function `gen_swissroll(n, p, w)`, where $n$ is the number of points, $p$ is the total number of dimensions, and $w$ is the vertical range in the third dimension. 
The first three dimensions form the classic $3\text{-}D$ Swiss roll shape. The $X_1 = t \cos(t)$, $X_2 = t \sin(t)$, $X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)$. For $p > 3$, the remaining $p - 3$ dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

### 2025-06-24 paper-cardinalR.tex
Updated cone data generation section

This function generates \texttt{n} points in a \texttt{p}-dimensional space, where the last dimension (\texttt{X\_p}) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip.

Points along the height axis are drawn from an exponential distribution to increase density near the tip:

\[
X_p = h_i \sim \text{Exp}(\lambda = 2/h), \quad \text{truncated at } h
\]

The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter:

\[
r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) \cdot \frac{h_i}{h}, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1
\]

@@

The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\). This classical non-orientable surface loops back on itself with a half-twist.

@@

To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape. The \(X_1 = t \cos(t), \quad X_2 = t \sin(t), \quad X_3 \sim U(w_1, w_2), \quad \text{with } t \sim U(0, 3\pi)\). For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.

>>>>

This function generates \(n\) points in a \(p\text{-}D\), where the last dimension (\(X_p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip. Points along the height axis are drawn from an exponential distribution to increase density near the tip. The \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\). The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter \(r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1\).

@@

The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\) structure. This classical non-orientable surface loops back on itself with a half-twist.

@@

To further generalize the Swiss roll structure and introduce realistic noise, we define a function \texttt{gen\_swissroll(n,\ p,\ w)}, where \(n\) is the number of points, \(p\) is the total number of dimensions, and \(w\) is the vertical range in the third dimension.
The first three dimensions form the classic \(3\text{-}D\) Swiss roll shape. The \(X_1 = t \cos(t)\), \(X_2 = t \sin(t)\), \(X_3 \sim U(w_1, w_2), \text{where } t \sim U(0, 3\pi)\). For \(p > 3\), the remaining \(p - 3\) dimensions are filled with small Gaussian noise to simulate high-dimensional complexity.
